---
title: "Longing for Chico - Part I"
output:
html_document: default
html_notebook: default
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE, 
                      echo = TRUE, dpi = 180)
options(width=80, dplyr.width = 150)
library(ggplot2)
library(scales)
theme_set(theme_minimal())
```

>*A marcha alegre se espalhou na avenida e insistiu*  
*A lua cheia que vivia escondida surgiu*  
*Minha cidade toda se enfeitou*  
*Pra ver a banda passar cantando coisas de amor*

One of my favorite artists is the Brazilian singer and songwriter Chico Buarque de Hollanda.  Though I have no real authority on the language (a few semesters in college), I have always found his songs exceptionally lyrical and poetic.  In general, Portugese vowels and the syntactic rhythms lend themselves particularly well to many Brazilian musical styles, particularly Samba and Bossa Nova.

<iframe width="560" height="315" frameborder="0" allowfullscreen src="https://www.youtube.com/watch?v=V7BaT6UKzog"></iframe>

In an effort to learn more about natural language processing, I decided to start learning by mining texts that I love - Samba and Bossa Nova songs.  However, I would have to shift what I am learning in text mining from English to Portuguese.  With David Robinson and Julia Silge's [*Tidy Text Mining with R*](http://tidytextmining.com/) in hand, I set out to learn a bit more about the inherent patterns of Chico's lyrics and understand his key themes and poetry.

First up, getting the data to begin with!

## Scraping Lyrics

The first step was, of course, to the lyrics of all of his songs.  I started at the website [letras.com](https://www.letras.com), which seemed to have a fairly comprehensive list of Chico's songs.  An entirely separate project would be corroborating this website, determining which (if any) songs and albums are missing, and filling those in. 

In the past, I usually use Beautiful Soup to scrape things in Python, but I wanted to see how easy it was to scrape data in R using [rvest](https://github.com/hadley/rvest).  Conclusion: very easy.

```{r letras_website}
## Libraries
library(tidyverse)
library(rvest)

## Site
site <- 'https://www.letras.com'
artist_page <- 'https://www.letras.com/chico-buarque/'
```

All I needed to do was scrape the title of every song listed on the artist page.  Each song title links to a separate page with the full lyrics for that song, so as I scraped the titles, I would also get the hyperlink to each individual page **html_chilren()**.

```{r scrape_song_list}
song_index <- artist_page %>%
  read_html() %>%
  html_nodes("#cnt-artist-songlist > div.cnt-list--alp > ul") %>%
  html_children()
```

I created two functions, get_lyrics() - for reading the song lyrics from each song's page and saving those lyrics to a list - and song_info() - a wrapper function to apply get_lyrics() to each song saved in my song_index object.


```{r main_functions}
get_lyrics <- function(link){
  scrape <- link %>% 
    read_html()
  lyrics <- scrape %>%
    html_nodes("#js-lyric-cnt > div.g-pr.g-sp > div.cnt-letra > article") %>%
    html_children() %>%
    gsub(pattern = '<br>', replacement = "\n") %>%
    gsub(pattern = '<.*?>', replacement = '')
  lyrics1 <- strsplit(lyrics,'\n')
  return(lyrics1)
}

song_info <- function(a){
  d <- a %>% html_children()
  song <- list(name=html_text(d))
  song[['link']] <- paste0(site,html_attr(d,"href"))
  song[['lyrics']] <- tryCatch(get_lyrics(song$link), error = function(e) e)
  return(song)
}
```


Then, all I needed to do was apply song_info() to each item in song_index, which I did using map() from the purrr package.

For close to 500 links, the script ran fairly quickly.  I saved each song as a list item, then saved that list to a **.rds** file so I wouldn't have to run it again. 

```{r put_it_all_together, eval=FALSE}
all_songs <- map(song_index,song_info)
names(all_songs) <- lapply(all_songs, function(x) x$name)
saveRDS(all_songs, file = 'chico_buarque_song_list.rds')
`````

Songs are structured as lines grouped by stanzas, and I wanted to preserve the integrity of that structure, so a nested list was my best bet.

>(Stanza 1)   
*Estava à toa na vida*  
*O meu amor me chamou*  
*Pra ver a banda passar*  
*Cantando coisas de amor*  

>(Stanza 2)  
*A minha gente sofrida*  
*Despediu-se da dor*  
*Pra ver a banda passar*  
*Cantando coisas de amor*
...

```{r song_structure, eval=TRUE,echo=FALSE}
a_banda <- song_info(song_index[[2]])
a_banda$lyrics[1:2]
`````

The script can easily be used to download all the songs for other artists on this site, which hopefully I'll use for another analysis (Joao Gilberto, Cartola, and Gilberto Gil - I'm comin' at ya!).




