---
title: "Longing for Chico - Part II"
output:
  html_document:
    theme: paper
    highlight: textmate
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE, 
                      echo = TRUE, dpi = 180)
options(width=80, dplyr.width = 150)
library(ggplot2)
library(scales)
library(elisalib)
theme_set(theme_ebvm())
library(tidyverse)
library(forcats)
```

Portuguese has a wonderful word: [*saudade*](https://en.wikipedia.org/wiki/Saudade).  While it doesn't have an English equivalent, it is usually translated as a longing for something in the past or something unattainable.  It's a beautiful sentiment that is often seen in literature and music, not to mention it's a wonderful word to say: *SOW-dahhd-gee*.  I decided that analyzing and quantifying the use of *saudade* in Chico Buarque's lyrics, and eventually other artists, would be a nice framework for my first foray into text mining.  So I tried to frame my exploration and analysis to see what sentiments, particularly *saudades*, drive the music of Chico Buarqe de Hollanda. 

Once I [got all the lyrics downloaded](previous blog post) for Chico Buarque,  I needed to do a little bit of cleaning of the data.  For starters, not all of the songs listed on the site are in Portuguese.  In fact, around 1970, Chico lived in exile in Italy after having been thrown in prison for *Roda Viva*, and has produced several albums in Italian (often translations of his Portuguese songs).  There were also several Spanish and French songs in the mix that I needed to filter.  I really only wanted to keep this analysis in one language!

Secondly, Chico Buarque has an extensive discography.  His first album came out in 1966, and he's been a prolific songwriter since then, so I thought it would be interesting to try to map each song to its original album, using the [discography](https://pt.wikipedia.org/wiki/Chico_Buarque) list on wikipedia, so I [created](clean_chico.R) a dataframe for each song in the song_list, matching up the songs to the wikipedia discography.  I still had a fair number of songs that were't perfectly matched, but I thought this was a fairly good start.

```{r album_df,eval=T,echo=F}
album_df <- read_csv(file = 'chico_buarque_album_df.csv')
album_df %>% head()

```

## Tidy Text Basics

I worked my way through several topics in [*Tidy Text Mining with R*](http://tidytextmining.com/), starting with word counts and text tokens.  I searched online for a list of Portugese [stop words](https://gist.github.com/alopes/5358189), which I saved to my local directory and simply read in as a table, and added 'pra' to my list.  *Para* is already a stop word, but it's contraction *'pra'* was not in the list, and I discovered that his songs make use of that word a lot.

I also created a few functions for manipulating the song lyrics into tidy data frames, which I stored in this [file](tidy_lyrics_helpers.R) and source along with the data that I read in.

```{r read_data, eval = T, echo = T}
source('tidy_lyrics_helpers.R')

all_songs <- readRDS(file = 'chico_buarque_song_list.rds')
album_df <- read_csv(file = 'chico_buarque_album_df.csv')
stop_words_pt <- read_table('stopwords.txt',col_names='word') %>%
  full_join(data_frame(word='pra'))
```


## A Banda

The first song on Chico Buarque's first album ('Chico Buarque de Hollanda', 1966) is called 'A Banda.'  He won the Viola de Ouro award at the 1966 Música Popular Brasileira Festival for this song, which can be seen here:

<iframe width="560" height="315" frameborder="0" allowfullscreen src="https://www.youtube.com/watch?v=ml8OpKjh4Ic"></iframe>

It's a lovely song about a small town where everyone - from the greedy miser to the ugly girl to the hidden moon - stops whatever causes causes their unhappiness to watch a band pass by singing about love.

We can start by selecting 'A Banda' from our all_songs list and converting it to a dataframe, a_banda_df.

Let's pull out each word in a lyric as a token using the unnest_tokens() function from the tidy text package, then we can remove the stop words using anti_join().  I also am simplifying the text so that repitition (i.e. multiple choruses) are removed.

```{r a_banda, eval = T, echo = T}
library(tidytext)
song <- all_songs[['A Banda']]

a_banda <- lyrics_df(song,simplify=T)


a_banda_words <- a_banda %>%
  unnest_tokens(word,Lyric) %>%
  anti_join(stop_words_pt)
a_banda_words
```

We can visualize the top words.  Unsurprisingly, *banda* is a top word, with vivia ('lived'), ver ('to see'), passar ('to pass'), parou ('stood'), contava ('told'), and amor ('love'), the next top words.  There really isn't enough data in this song to take these word counts so seriously, but if we look at a larger body of his works, we'll hopefully be able to see more patterns emerge.

```{r top_words, eval = T, echo = T}
a_banda_words %>%
  count(word, sort=TRUE) %>%
  mutate(word = reorder(word,n)) %>%
  filter(n>1) %>%
  ggplot(aes(x=word, y=n,fill=n)) +
  geom_col()+
  scale_fill_ebvmCont(palette='Ordem')+
  xlab(NULL)+
  coord_flip()+
  labs(title = 'A Banda')
```


## Chico Buarque de Hollanda, Vol. 1

Let's look at Chico Buarque's entire first album, the self-titled *Chico Buarque de Hollanda*, which came out in 1966.  We can actually look at the top word count both when we simplify the songs (remove repeated stanzas) and when we keep the repetition.

```{r vol1, eval = T, echo = T}
vol1_songs <- all_songs[album_df %>%
  filter(Album == 'Chico Buarque de Hollanda' & Year == 1966) %>%
  collect %>%
  .[['Song']]]

tidy_count <- clean_lyrics(vol1_songs,simp=FALSE) %>%
  unnest_tokens(word,Lyric) %>%
  anti_join(stop_words_pt) %>%
  count(word,sort=TRUE) %>%
  mutate(word = reorder(word, n)) %>%
  mutate(Simplified = F)

simplified_count <- clean_lyrics(vol1_songs,simp=TRUE) %>%
  unnest_tokens(word,Lyric) %>%
  anti_join(stop_words_pt) %>%
  count(word,sort=TRUE) %>%
  mutate(word = reorder(word, n)) %>%
  mutate(Simplified = T)

all_counts <- full_join(tidy_count,simplified_count)

all_counts %>% 
  group_by(Simplified) %>%
  top_n(20,n) %>%
  ungroup() %>%
  mutate(word = fct_reorder(as.factor(word),n)) %>%
  ggplot(aes(x=word, y = n, fill=Simplified)) +
  geom_col(position='dodge') + 
  scale_fill_ebvmDisc(palette='Ordem')+
  xlab(NULL) +
  coord_flip() +
  labs(title = 'Chico Buarque de Hollanda, Vol. 1',
       subtitle = 'Album Word Count')

```


Let's reconsider the top word count just looking at the simplified songs.  The top words used throughout the album are Samba and Esperando (hoping/waiting).

```{r vol1_simplified, echo = F, eval = T}

all_counts %>% 
  filter(Simplified == T) %>%
  top_n(20,n) %>%
  mutate(word = fct_reorder(as.factor(word),n)) %>%
  ggplot(aes(x=word, y = n, fill=n)) +
  geom_col(position='dodge') + 
  scale_fill_ebvmCont(palette='Ordem')+
  xlab(NULL) +
  coord_flip() +
  labs(title = 'Chico Buarque de Hollanda, Vol. 1',
       subtitle = 'Album Word Count (Simplified Songs)')

```

Was samba always a highly used word?  What about my hypothesis of saudade?  And esperando?  If we look at a trend line of these words, its fairly disjointed.  I organized the word by rank (word count in that year), and unfortunately, there isn't necessarily one distinct pattern.

```{r word_evolution, echo = F, eval = T}

all_df <- clean_lyrics(all_songs[album_df$Song], simp = TRUE) %>%
  left_join(album_df)

year_count <- all_df %>%
  unnest_tokens(word,Lyric) %>%
  anti_join(stop_words_pt) %>%
  group_by(Year) %>%
  na.omit() %>%
  count(word, sort=T) %>%
  ungroup() %>%
  group_by(Year) %>%
  arrange(-n) %>%
  mutate(Rank = row_number()) %>%
  ungroup()
year_count %>%
  filter(word == 'saudade')

year_count %>%
  filter(word %in%c('samba','saudade','esperando')) %>%
  mutate(Rank = as.integer(Rank)) %>%
  ggplot(aes(x = Year, y = Rank, color=word, group = word)) +
  geom_line()+
  scale_y_reverse() +
  geom_point()

```

Interestingly, esperando drops off the chart in the early 1970s.  Samba remains relatively popular through the 1980s with a recent resurgence (Chico Buarque seems to have written less original songs in later years).  Finally, *saudade* is not ranked that highly but for two times, in 1978 and 1983.

## Conclusion

So where were the saudades that I thought I'd be finding throughout Chico Buarque's lyrics?